---
title: "Covid91 vaccine study Final2024"
author: "Maximiliano Ortiz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
males <- FinalData %>% filter(sex == "M")
nrow(males)
male_test <- prop.test(sum(males$infected == 1), nrow(males), p = 0.5)
male_test
```

The results for males were also highly significant, meaning the vaccine clearly helps reduce infections in this group. This supports its effectiveness for both men and women.

## Females

```{r}
females <- FinalData %>% filter(sex == "F")
nrow(females)
female_test <- prop.test(sum(females$infected == 1), nrow(females), p = 0.5)
female_test
```

The proportion test for females indicates that the observed infection rate is significantly lower than 50%. The extremely small p-value provides strong evidence that the vaccine offers a protective effect for female participants, reducing infection risk beyond what would be expected by chance.

## LGBTQ

```{r}
lgbtq <- FinalData %>% filter(LGBTQ == "gay")
nrow(lgbtq)
lgbtq_test <- prop.test(sum(lgbtq$infected == 1), nrow(lgbtq), p = 0.5)
lgbtq_test
```

The analysis of LGBTQ participants indicates a significant protective effect of the vaccine, evidenced by a p-value well below 0.05. These results show that vaccination substantially reduces infection risk within this group, supporting the vaccine’s effectiveness across diverse sexual orientations.

## Druggies

```{r}
druggies <- FinalData %>% filter(DrugUser == "yes")
nrow(druggies)
druggies_test <- prop.test(sum(druggies$infected == 1), nrow(druggies), p = 0.5)
druggies_test
```

The proportion test for participants who use drugs is also statistically significant, indicating that the vaccine effectively reduces infection rates within this group. This finding is especially meaningful given the elevated health risks often faced by drug users and further supports the vaccine’s efficacy across diverse populations.

# Overall Results and Conclusions

In summary, the vaccine shows statistically significant effectiveness across all examined subgroups females, males, LGBTQ participants, and individuals who use drugs. In each case, infection rates are significantly below 50%, demonstrating broad protective effects across diverse demographic and behavioral groups. These results provide strong evidence of the vaccine’s overall efficacy within the study population and underscore its potential for widespread protective benefits.
